CRUSE to Safe Cycling in Ireland

### 2023-10-09 paper.qmd
Comment out TODO
Shorten abstract

# # Uncomment for draft versions:
format: 
  html: default
  docx: default
# # Uncomment for formatted version
#   elsevier-pdf:
#     keep-tex: true
#     journal:
#       name: Journal Name
#       formatting: preprint
#       model: 3p
#       cite-style: super
# # #   # pdf: default

# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, EXETER, EX5 1FY, UK
  - name: "Gary O'Tool"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  Under the EU Road infrastructure safety management (RISM) directive, the National Road Safety Strategy (RSS) and the Climate Action Plan, Transport Infrastructure Ireland (TII) has a remit for road safety and decarbonizing a predominantly road-based network in Ireland.

  The Cycle Route Uptake and Scenario Estimation (CRUSE) Tool was commissioned by TII and developed by the Leeds Institute for Transport Studies (ITS) and AECOM To address data needs for safety, business case development and project evaluation on the National Road Network (NRN).
  Cycling in Ireland represents only 3% of total modal share as of the latest Census (2016), but accounts for 20% of serious injuries and 7% of all fatalities.
  High resolution datasets on baseline and potential cycling levels are needed to meet Ireland's climate and safety targets.

  The CRUSE project builds on the methods underlying the Propensity to Cycle Tool for England and Wales (PCT), which had a major impact on strategic cycle network planning in the UK.
  Like the PCT, CRUSE is based on open source software and reproducible data science. The resulting web application is available at <https://cruse.bike/>, enabling planners, engineers, and other stakeholders to make more evidence-based decisions for the NRN.
  CRUSE goes beyond the PCT in several important ways, modeling denser networks at higher levels of geographic resolution and additional purposes, including simulated recreational, social, shopping and personal trips, in addition to trips to work and primary, secondary and tertiary education locations.
  The tool provides key 'denominator' information to enable reporting of collision rates per billion km cycled.

  Another key feature of CRUSE is its use of Open Street Map (OSM) data, which is continuously updated, and used by [Cyclestreets.net](https://www.cyclestreets.net/) for route calculation and "cycle friendliness" estimates of all links on the network.
  CRUSE provides three network types: the 'Fastest' network highlights routes for directness; 'Quietest' prioritises cycle friendliness, and 'Balanced' represents a balance between speed and comfort.
  Trips are generated based on origin and destination data from the 2016 Census in combination with modeled demand data to estimate current and potential future cycling levels at area, route, and network levels for each county in Ireland.

  For investment in cycling to deliver on the huge potential benefits, cycling must become safer.
  With growth in the E-bike market, CRUSE will help inform inter-urban and rural networks to support the transfer of trips to sustainable modes for longer journeys.
  Provision of open access data in a publicly available web tool will address the challenges faced by National Road Authorities (NRAs) across Ireland in meeting their sustainability commitments.
  The experience of developing a deploying CRUSE are relevant to countries with sustainable transport policies worldwide.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
date: last-modified
---

@@

TODO: Add description of route types in Naas based on local knowledge and @fig-route-types.

>>>>

# # Uncomment for draft versions:
format: 
  # html: default
  # docx: default
# Uncomment for formatted version
  elsevier-pdf:
    linestretch: 2
    keep-tex: true
    journal:
      name: European Transport Research Review
      formatting: preprint
      model: 3p
      cite-style: super
# # #   # pdf: default

# # Uncomment the following line to produce the cover page
author:
  - name: Robin Lovelace
    email: r.lovelace@leeds.ac.uk
    corresponding: true
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Joey Talbot
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Eugeni Vidal-Tortosa
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Hussein Mahfouz
    affiliations:
      - name: University of Leeds
        address: Institute for Transport Studies, Leeds, LS2 9JT, United Kingdom
  - name: Elaine Brick
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Peter Wright
    affiliations:
      - name: AECOM
        address: Winslade House, Winslade Park, Manor Drive, Clyst St Mary, EXETER, EX5 1FY, UK
  - name: "Gary O'Tool"
    affiliations:
      - name: AECOM
        address: Unit 6, Galway Technology Park, Parkmore, Galway, Ireland
  - name: Dan Brennan
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
  - name: Suzanne Meade
    affiliations:
      - name: Transport Infrastructure Ireland
        address: Parkgate Business Centre, Parkgate Street, Dublin 8, D08 DK10, Ireland
abstract: |
  Legislation mandates improved road safety and decarbonisation of predominantly road-based transport network in Ireland and other countries. The Cycle Route Uptake and Scenario Estimation (CRUSE) Tool was commissioned to support cycle network planning, safety interventions, business case development and project evaluation.
  Cycling in Ireland represents only 3% of total modal share as of the latest Census (2016), but accounts for 20% of serious injuries and 7% of all fatalities.
  High resolution datasets on baseline and potential cycling levels are needed to meet Ireland's climate and safety targets.

  The CRUSE Tool is based on open source software and reproducible data science. The resulting web application is available at <https://cruse.bike/>, enabling planners, engineers, and other stakeholders to make more evidence-based decisions.
  CRUSE goes beyond previous work in several important ways.
  It models networks at high levels of geographic resolution, simulates a wide range of trips including recreational, social, shopping and personal trips, in addition to trips to work and primary, secondary and tertiary education locations.
  The tool provides key 'denominator' information to enable reporting of collision rates per billion km cycled.

  Another key feature of CRUSE is its use of frequently-updated OpenStreetMap data for route calculation and "cycle friendliness" estimates of all links on the network.
  CRUSE provides three network types: the 'Fastest' network highlights routes for directness; 'Quietest' prioritises cycle friendliness, and 'Balanced' represents a balance between speed and comfort.
  Trips are generated based on origin and destination data from the 2016 Census in combination with modeled demand data to estimate current and potential future cycling levels at area, route, and network levels for each county in Ireland.

  For investment in cycling to deliver on the huge potential benefits, cycling must become safer.
  With growth in the E-bike market, CRUSE will help inform inter-urban and rural networks to support the transfer of trips to sustainable modes for longer journeys.
  Provision of open access data in a publicly available web tool will address the challenges faced by counties across Ireland in meeting their sustainability commitments.
  The experience of developing a deploying CRUSE are relevant to countries with sustainable transport policies worldwide.
keywords: 
  - Cycling
  - Open-source
  - Road Safety
  - Active Travel
date: last-modified
---

@@

<!-- TODO: Add description of route types in Naas based on local knowledge and @fig-route-types. -->
